Drought and its Effects on Economy

### 2018-11-16 presentations/dec2017BMEcSeminar/presentation3.tex
adding files, initial

\large{\hspace{1cm}\textbf{Effects of droughts on economy in Kenya}}
\end{frame}

@@

@@

\note{Heim = a review of the 20tieth century indices, Trenberth = IPCC report, summary for policy makers \\
\vspace{0.3cm}f  \\
\vspace{0.3cm} f\\
\vspace{0.3cm}

@@

\note{On short timescales, the SPI is closely related to soil moisture, while at longer timescales, the SPI can be related to groundwater and reservoir storage.\\
\vspace{0.3cm}SPI: usuallygamma or a Pearson Type III distribution \\
\vspace{0.3cm} can be used across regions with very different climates\\
\vspace{0.3cm}SPEI no assumptions about underlining system
\\
\vspace{0.3cm} requires more data than SPI
\\
\vspace{0.3cm} SPEI sensitive to method that calculate potential evapotranspiration

\\
\vspace{0.3cm}
Evapotranspiration (ET) is the sum of evaporation and plant transpiration from the Earth's land and ocean surface to the atmosphere. Evaporation accounts for the movement of water to the air from sources such as the soil, canopy interception, and waterbodies. 
}

@@

\frametitle{Data}\label{data}

mozna taky trochu popsat data?? nebo az u napadu, co budu delat??

@@

\note{\\
\vspace{0.3cm}A CGE model consists of equations describing model variables and a database (usually very detailed) consistent with these model equations\\
\vspace{0.3cm}a\\
\vspace{0.3cm}
\vspace{0.3cm}
}

@@

\note{
\vspace{0.3cm}the second study: robinson and willenbockel: effects of the 20 percent reductions caused by drought cause 2.3 percent decrease in Ethipia GDP. Also models price transition between international and domestic prices..
\\
\vspace{0.3cm}robinson: also model food aid (decrease prices of food>>good for all net food purchasers
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}

@@

\end{footnotesize}
\end{itemize}
\end{block}
\end{frame}
\note{\\

\vspace{0.3cm}
\vspace{0.3cm}a
\\
\vspace{0.3cm}a
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}

@@

\end{frame}
\note{Lesk: 9.1 percent reduction is attributable to extreme heat.
They did not find any significant effect of extreme cold and 
oods on production.

\vspace{0.3cm} mehrabi: estimate counter factual production in disaster years
\vspace{0.3cm}mehrabi:units are a bit like a panel - years x commodity x country x disaster (disaster years and not disaster years
\\mehrabi: difference in production of disaster years and non-disaster year average. Cumulative impact: sum per countries and disasters. Loss relatively to (divided by) production + loss of production
\vspace{0.3cm}mehrabi: the most severe losses angola, botswana, USA, Australie, Paraguay, Angola
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}

@@

\item Effects on production (production function approach)
\begin{itemize}
\item \textbf{\cite{Ochieng2016}}
\begin{itemize}
\item Effect of climate variability and change on crop revenue
\item Augmented production function, Kenya
\item  Household FE
\end{itemize}
\item \textbf{\cite{Deschenes2007Ric}:} US, positive effect of global warming
\end{itemize}
\item Effects on land prices or yields/acre (Ricardian analysis)
\begin{itemize}
\item  \textbf{\citet{kabubo2007}:} Kenya, warming harmful

@@

\vspace{0.3cm}brown, Khsingar: developing world (Africa, South Asia, Latin America)	
\\
\vspace{0.3cm}Seo et. al: Africa- 16 agro-ecological zones
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}
% tady mozna pridat nejake ricardian studie a rozdelit to na dva slidy

\section{My Idea}

@@

\frametitle{My idea - panel estimation}\label{MySuggestion}

@@

\item \textbf{Outcome variable}

\begin{itemize}
\item Volumes of production (crop specific, total)
\item Alternatively food prices, yields/per acre
\end{itemize}

\item \textbf{Units of analysis}
\begin{itemize}
\item Counties in Kenya $\times$ year
\
\end{itemize}

@@

\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
}

@@

\frametitle{My idea - panel estimation}\label{MySuggestion}

@@

\frametitle{My idea - panel estimation}\label{MySuggestion}

@@

\item \textbf{Control variables ?}

\begin{itemize}
\item GDP, population, land area, soil characteristics, climate (2 seasons)?
\end{itemize}

@@

\frametitle{Cumulative logits model}\label{Results-Logit2}

 $Y$ is a categorical response with $M$ categories $(m=1...M)$

\begin{equation}\label{PCumLog}
\begin{array}{lclcl}
logit  \big[P( Y \leq m| x) \big] &=& \alpha_m + \boldsymbol{\beta'x}, ~~m=1,...,M-1
\end{array}
\end{equation}

\begin{align*}\label{Stickin}
\begin{array}{lcl}
\text{Substitute estimates into:} \hspace{1cm}\hat{P}(Y\leq m) &=& \frac{exp(\hat{\alpha_m}+ \hat{\beta} x)}{1+exp(\hat{\alpha_m}+ \hat{\beta} x)}
\end{array}
\end{align*}

\vspace{0.5cm}
\begin{itemize}
\item \textbf{Proportional odds assumption}
\begin{itemize}
\item Likelihood ratio test for comparison with multinomial logit insignificant $\boldsymbol{\Rightarrow}$ assumption satisfied

@@

\section*{~~}

\begin{frame}
\centering
\huge{Thank you for attention}
\end{frame}

>>>>

\large{\hspace{1cm}\textbf{Effects of droughts on economy in Kenya}}
\end{frame}
\note{A}

@@

\note{A}

@@

\note{Heim = a review of the 20tieth century indices, Trenberth = IPCC report, summary for policy makers \\

@@

\note{On short timescales, the SPI is closely related to soil moisture, while at longer timescales, the SPI can be related to groundwater and reservoir storage.\\
\vspace{0.3cm}SPI: usuallygamma or a Pearson Type III distribution \\
\vspace{0.3cm} can be used across regions with very different climates\\
\vspace{0.3cm}SPEI no assumptions about underlining system
\\
\vspace{0.3cm} requires more data than SPI
\\
\vspace{0.3cm} SPEI sensitive to method that calculate potential evapotranspiration
\vspace{0.3cm}
Evapotranspiration (ET) is the sum of evaporation and plant transpiration from the Earth's land and ocean surface to the atmosphere. Evaporation accounts for the movement of water to the air from sources such as the soil, canopy interception, and waterbodies. 
}

@@

@@

\note{
\vspace{0.3cm}A CGE model consists of equations describing model variables and a database (usually very detailed) consistent with these model equations\\
\vspace{0.3cm}
\vspace{0.3cm}
}

@@

\note{
\vspace{0.3cm}the second study: robinson and willenbockel: effects of the 20 percent reductions caused by drought cause 2.3 percent decrease in Ethipia GDP. Also models price transition between international and domestic prices..
\\
\vspace{0.3cm}robinson: also model food aid (decrease prices of food>>good for all net food purchasers
\\qe
\vspace{0.3cm}a\\}

@@

\end{footnotesize}
\end{itemize}
\end{block}
\end{frame}
\note{

\vspace{0.3cm}
\vspace{0.3cm}a
\\
\vspace{0.3cm}a
\\
\vspace{0.3cm}a
}

@@

\end{frame}
\note{Lesk: 9.1 percent reduction is attributable to extreme heat.
They did not find any significant effect of extreme cold and 
oods on production.
\vspace{0.3cm} mehrabi: estimate counter factual production in disaster years
\vspace{0.3cm}mehrabi:units are a bit like a panel - years x commodity x country x disaster (disaster years and not disaster years
\\mehrabi: difference in production of disaster years and non-disaster year average. Cumulative impact: sum per countries and disasters. Loss relatively to (divided by) production + loss of production
\vspace{0.3cm}mehrabi: the most severe losses angola, botswana, USA, Australie, Paraguay, Angola\\
\vspace{0.3cm}a
\vspace{0.3cm}
}

@@

\item Effects on production (production function approach)
\begin{itemize}
\item \textbf{\cite{Ochieng2016}}
\begin{itemize}
\item Effect of climate variability and change on crop revenue
\item Augmented production function, Kenya
\item  Household FE
\end{itemize}
\item \textbf{\cite{Deschenes2007Ric}:} US county level, positive effect of global warming
\end{itemize}
\item Effects on land prices or yields/acre (Ricardian analysis)
\begin{itemize}
\item  \textbf{\citet{kabubo2007}:} Kenya, warming harmful

@@

\vspace{0.3cm}brown, Khsingar: developing world (Africa, South Asia, Latin America)	\\
\vspace{0.3cm}Seo et. al: Africa- 16 agro-ecological zones\\
\vspace{0.3cm}a\\
\vspace{0.3cm}}
% tady mozna pridat nejake ricardian studie a rozdelit to na dva slidy

\section{My Suggestion}

@@

\frametitle{My suggestion - panel estimation}\label{MySuggestion1}

@@

\item \textbf{Response variable}

\begin{itemize}
\item Volumes of production (crop specific, total)
\item Profit per acre \normalsize{\citep{Deschenes2007Ric}}

\begin{itemize}
\item \footnotesize{(Value of agricultural products - prod. expanses)/acres (crops, pasture, grazing)}\\ 
\end{itemize}
\end{itemize}
\item \textbf{Units of analysis}
\begin{itemize}
\item Counties in Kenya $\times$ year
\
\end{itemize}

@@

\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county\\
\vspace{0.3cm}a\\
\vspace{0.3cm}}

@@

\frametitle{My suggestion - panel estimation}\label{MySuggestion2}

@@

\frametitle{My suggestion - panel estimation}\label{MySuggestion3}

@@

\item \textbf{Control variables ? - subject to availability}

\begin{itemize}
\item GDP, soil quality data, population,average land area, average value assets,  climate, degree days?
\end{itemize}

@@

\frametitle{My idea - panel estimation}\label{MySuggestion4}

\begin{align*}
\begin{array}{lcl}
 Y_{i,t} &=& \alpha_i +\gamma_t + \delta D'_{i,t} +  \boldsymbol{\beta X_{i,t}}+ \epsilon_{i,t}
\end{array}
\end{align*}

\begin{table}
\begin{small}
\begin{tabular}{lll}

$Y_{i,t}$ &$=$& Response variable (food production/price), county $i$ in year $t$\\
$\alpha_i$ &$=$& Fixed effects, county $i$\\
{\color{darkblue}$\delta$ }&{\color{darkblue}$=$}&{\color{darkblue}{Effect of drought on economy}}\\
$D_{i,t}$ &$=$& Indicator variable\\
& &  $D=1$ if drought in county $i$ in year $t$, $D=0$ otherwise\\
$\boldsymbol{\beta}$ &$=$& Vector of effects of other covariates\\
$X_{i,t}$ &$=$& Matrix of values of other covariates in county $i$ in year $t$\\
$\epsilon_{i,t}$ &$=$& Error term \\
{\color{DarkGreen}$\gamma_t$} & {\color{DarkGreen}$=$}& {\color{DarkGreen}Year specific indicator?}
\end{tabular}
\end{small}
\end{table}

@@

\section*{}

\frame[allowframebreaks]{ 
\centering
\huge{Thank you for attention}
}

### 2018-11-16 presentations/dec2017BMEcSeminar/presentation4.tex
adding files, initial

\item \textbf{Conceptual definitions:} dictionary types, usually defining boundaries of the concept of drought \\ 
e.g. \textit{An extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region \citep{schneider1996}}
\item \textbf{Operational definitions:} Foundation for an effective early warning system \\ e.g. SPI, PDSI

@@

\item Significant milestone in history of drought severity quantification
\item Based on a hydrological accounting system
\item Incorporate antecedent precipitation, moisture supply and moisture demand
\end{itemize}
\item \textbf{Standardized Precipitation Index} (SPI, \citealp{SPI})
\begin{itemize}
\item Can be interpreted as the number of standard deviations by which the observed value differs from the long-term mean
\item Standardised departure of observed precipitation from a chosen probability distribution function which models the precipitation data \citep{SPIonline}

@@

@@

\begin{itemize}
\item \underline{\textbf{\cite{robinson2010}}}
\begin{itemize}
\item Ethiopia, Social Accounting Matrix (SAM)

\item Model drought as a $20\%$ reduction in crop productivity and $20$ reduction of livestock capital
\item Effects of shocks in production on cereal prices and food consumption
\item Includes a scenario with large-scaled inflow of wheat financed by rest of the world
\end{itemize}
\end{itemize}
\end{block}
\end{frame}

\item The simulation use a 'balanced' macro closure in which aggregate investment, government demand, and consumption are fixed shares of total absorption
\item Intermediate inputs into production are determined as fixed shares
of the quantity of output

@@

\begin{itemize}
\item National production losses per disaster (droughts, floods and extreme temperatures)
\item Worldwide, superposed epoch analysis
\item On average $10.1\%$ reduction of cereal production can be linked to drought
\end{itemize}
\item \textbf{\cite{Mehrabi2017}}
\begin{itemize}
\item Cumulative production losses linked to extreme heat and drought events
\item Per years (53), countries (131), commodities (6)
\end{itemize}

>>>>

\item \textbf{Conceptual definitions:} dictionary types, usually defining boundaries of the concept of drought \\ 
e.g. \textit{An extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region \citep{schneider1996}}
\item \textbf{Operational definitions:} Foundation for an effective early warning system \begin{itemize}
\item e.g.  Standardized Precipitation Index (SPI), Palmer Drought Severity Index (PDSI)
\end{itemize} 
\end{itemize}

@@

\item Significant milestone in history of drought severity quantification
\item Based on a hydrological accounting system
\item Incorporate antecedent precipitation, moisture supply and moisture demand
\end{itemize}
\item \textbf{Standardized Precipitation Index} (SPI, \citealp{SPI})
\begin{itemize}
\item Can be interpreted as the number of standard deviations by which the observed value differs from the long-term mean
\item Standardized departure of observed precipitation from a chosen probability distribution function which models the precipitation data \citep{SPIonline}

@@

\vspace{0.3cm}

\item \underline{\textbf{\cite{robinson2010}}}
\begin{itemize}
\item Ethiopia, Social Accounting Matrix (SAM)

\item Model drought as a $20\%$ reduction in crop productivity and $20\%$ reduction of livestock capital
\item Effects of shocks in production on cereal prices and food consumption
\item Results in $2.3\%$ decrease in total GDP
\item Includes a scenario with large-scaled inflow of wheat financed by rest of the world
\end{itemize}
\end{itemize}
\end{block}
\end{frame}

\item The simulation use a 'balanced' macro closure in which aggregate \textbf{investment, government demand, and consumption are fixed shares of total absorption}
\item Intermediate inputs into production are determined as fixed shares
of the quantity of output

@@

@@

\begin{itemize}
\item National production losses per disaster (droughts, floods and extreme temperatures)
\item Worldwide, superposed epoch analysis
\item On average $10.1\%$ reduction of cereal production can be linked to drought
\end{itemize}
\item \textbf{\cite{Mehrabi2017}}
\begin{itemize}
\item Cumulative production losses linked to extreme heat and drought events
\item Per years (53), countries (131), commodities (6)
\item Most severe losses in Botswana ($5.7\%$), Paraguay ($5.5\%$), Nigeria ($4.8\%$)
\end{itemize}
